---
title: "trump_DataOverTime_EG"
author: "Evan Garcia"
date: "2026-03-01"
output: pdf_document
---

```{r}
library(dplyr)
library(lubridate)
library(ggplot2)
library(gridExtra)

# Load data
data <- read.csv("filtered_trump_dat.csv")

# Parse dates globally
data$datetime <- ymd_hms(data$date)
data$day <- as.Date(data$datetime)
```


## Section 1 - Candidacy Announcement Analysis (June 16, 2015)

We compare average likes (favorite_count), retweets (repost_count), and daily post counts in the month **before** vs. the month **after** Trump announced his candidacy.

- **Pre-announcement window:** May 16 – June 15, 2015
- **Post-announcement window:** June 16 – July 16, 2015

```{r}
# Parse dates
data$datetime <- ymd_hms(data$date)
data$day <- as.Date(data$datetime)
# 2. Add this line here to clean the data immediately
data <- data %>% filter(!is.na(day))
# Define windows
announce_date <- as.Date("2015-06-16")
pre_start  <- announce_date - 31   # May 16, 2015
pre_end    <- announce_date - 1    # June 15, 2015
post_start <- announce_date        # June 16, 2015
post_end   <- announce_date + 30   # July 16, 2015

# Split data
pre_data  <- data %>% filter(day >= pre_start & day <= pre_end)
post_data <- data %>% filter(day >= post_start & day <= post_end)
```

### Processing 

```{r}
# Compute daily post counts for each window
daily_pre <- pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

daily_post <- post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
all_pre_days  <- data.frame(day = seq(pre_start, pre_end, by = "day"))
all_post_days <- data.frame(day = seq(post_start, post_end, by = "day"))

daily_pre  <- left_join(all_pre_days, daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
daily_post <- left_join(all_post_days, daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_month <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(pre_data$favorite_count), mean(pre_data$repost_count), mean(daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(post_data$favorite_count), mean(post_data$repost_count), mean(daily_post$n_posts)), 2),
  p_value = c(
    t.test(pre_data$favorite_count, post_data$favorite_count)$p.value,
    t.test(pre_data$repost_count, post_data$repost_count)$p.value,
    t.test(daily_pre$n_posts, daily_post$n_posts)$p.value
  )
)
summary_month$Significant_0.05 <- ifelse(summary_month$p_value < 0.05, "Yes", "No")

# compute for year 
yr_pre_start  <- as.Date("2014-06-16")
yr_pre_end    <- as.Date("2015-06-15")
yr_post_start <- as.Date("2015-06-16")
yr_post_end   <- as.Date("2016-06-15")

yr_pre_data  <- data %>% filter(day >= yr_pre_start & day <= yr_pre_end)
yr_post_data <- data %>% filter(day >= yr_post_start & day <= yr_post_end)

yr_daily_pre <- yr_pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

yr_daily_post <- yr_post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
yr_all_pre_days  <- data.frame(day = seq(yr_pre_start, yr_pre_end, by = "day"))
yr_all_post_days <- data.frame(day = seq(yr_post_start, yr_post_end, by = "day"))

yr_daily_pre  <- left_join(yr_all_pre_days, yr_daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
yr_daily_post <- left_join(yr_all_post_days, yr_daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_year <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(yr_pre_data$favorite_count), mean(yr_pre_data$repost_count), mean(yr_daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(yr_post_data$favorite_count), mean(yr_post_data$repost_count), mean(yr_daily_post$n_posts)), 2),
  p_value = c(
    t.test(yr_pre_data$favorite_count, yr_post_data$favorite_count)$p.value,
    t.test(yr_pre_data$repost_count, yr_post_data$repost_count)$p.value,
    t.test(yr_daily_pre$n_posts, yr_daily_post$n_posts)$p.value
  )
)
summary_year$Significant_0.05 <- ifelse(summary_year$p_value < 0.05, "Yes", "No")
```

### Results 

```{r}
summary_month$Window <- "Month"
summary_year$Window  <- "Year"
combined <- rbind(summary_month, summary_year)
combined <- combined[, c("Window", "Metric", "Pre_Mean", "Post_Mean", "p_value", "Significant_0.05")]
combined
```

```{r, fig.width=10, fig.height=4}
# --- Bar Charts: Pre vs Post Announcement (Month) ---
# Reshape the summary_month table for easier plotting
library(tidyr)
plot_month <- summary_month %>%
  filter(Metric != "Daily Post Count") %>%
  select(Metric, Pre_Mean, Post_Mean) %>%
  pivot_longer(cols = c(Pre_Mean, Post_Mean), names_to = "period", values_to = "Mean") %>%
  mutate(period = factor(period, levels = c("Pre_Mean", "Post_Mean"), labels = c("Pre", "Post")))

# Plot Likes Mean
bar1 <- ggplot(plot_month %>% filter(Metric == "Likes (favorite_count)"), 
               aes(x = period, y = Mean, fill = period)) +
  geom_bar(stat = "identity", color = "black", width = 0.6) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Mean Likes: Pre vs Post (Month)", x = "", y = "Average Likes") +
  theme_minimal() + theme(legend.position = "none")

# Plot Retweets Mean
bar2 <- ggplot(plot_month %>% filter(Metric == "Retweets (repost_count)"), 
               aes(x = period, y = Mean, fill = period)) +
  geom_bar(stat = "identity", color = "black", width = 0.6) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  scale_y_continuous(labels = scales::comma) +
  labs(title = "Mean Retweets: Pre vs Post (Month)", x = "", y = "Average Retweets") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(bar1, bar2, ncol = 2)
```

---

## Year-Window Comparison (June 16, 2014–2015 vs. June 16, 2015–2016)

We now widen the comparison to the full **year before** vs. **year after** the announcement.

- **Pre-announcement window:** June 16, 2014 – June 15, 2015
- **Post-announcement window:** June 16, 2015 – June 15, 2016


### Plots — Candidacy Announcement (Year Window)

```{r, fig.width=10, fig.height=4}
bx3 <- ggplot(cand_year, aes(x = period, y = favorite_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Likes: Pre vs Post Announcement (Year)", x = "", y = "Likes") +
  theme_minimal() + theme(legend.position = "none")

bx4 <- ggplot(cand_year, aes(x = period, y = repost_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Retweets: Pre vs Post Announcement (Year)", x = "", y = "Retweets") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(bx3, bx4, ncol = 2)
```

---

## Section 2 - Election Loss Analysis (November 3, 2020)

We compare average likes (favorite_count), retweets (repost_count), and daily post counts in the month **before** vs. the month **after** Trump lost the 2020 election.

- **Pre-election window:** October 3 – November 2, 2020
- **Post-election window:** November 3 – December 3, 2020

```{r}
# Define election-loss windows
elec_date <- as.Date("2020-11-03")
elec_pre_start  <- elec_date - 31   # Oct 3, 2020
elec_pre_end    <- elec_date - 1     # Nov 2, 2020
elec_post_start <- elec_date         # Nov 3, 2020
elec_post_end   <- elec_date + 30    # Dec 3, 2020

# Split data
elec_pre_data  <- data %>% filter(day >= elec_pre_start & day <= elec_pre_end)
elec_post_data <- data %>% filter(day >= elec_post_start & day <= elec_post_end)
```

### Processing

```{r}
# Compute daily post counts for each window
elec_daily_pre <- elec_pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

elec_daily_post <- elec_post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
elec_all_pre_days  <- data.frame(day = seq(elec_pre_start, elec_pre_end, by = "day"))
elec_all_post_days <- data.frame(day = seq(elec_post_start, elec_post_end, by = "day"))

elec_daily_pre  <- left_join(elec_all_pre_days, elec_daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
elec_daily_post <- left_join(elec_all_post_days, elec_daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_elec_month <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(elec_pre_data$favorite_count), mean(elec_pre_data$repost_count), mean(elec_daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(elec_post_data$favorite_count), mean(elec_post_data$repost_count), mean(elec_daily_post$n_posts)), 2),
  p_value = c(
    t.test(elec_pre_data$favorite_count, elec_post_data$favorite_count)$p.value,
    t.test(elec_pre_data$repost_count, elec_post_data$repost_count)$p.value,
    t.test(elec_daily_pre$n_posts, elec_daily_post$n_posts)$p.value
  )
)
summary_elec_month$Significant_0.05 <- ifelse(summary_elec_month$p_value < 0.05, "Yes", "No")

# compute for year
elec_yr_pre_start  <- as.Date("2019-11-03")
elec_yr_pre_end    <- as.Date("2020-11-02")
elec_yr_post_start <- as.Date("2020-11-03")
elec_yr_post_end   <- as.Date("2021-11-02")

elec_yr_pre_data  <- data %>% filter(day >= elec_yr_pre_start & day <= elec_yr_pre_end)
elec_yr_post_data <- data %>% filter(day >= elec_yr_post_start & day <= elec_yr_post_end)

elec_yr_daily_pre <- elec_yr_pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

elec_yr_daily_post <- elec_yr_post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
elec_yr_all_pre_days  <- data.frame(day = seq(elec_yr_pre_start, elec_yr_pre_end, by = "day"))
elec_yr_all_post_days <- data.frame(day = seq(elec_yr_post_start, elec_yr_post_end, by = "day"))

elec_yr_daily_pre  <- left_join(elec_yr_all_pre_days, elec_yr_daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
elec_yr_daily_post <- left_join(elec_yr_all_post_days, elec_yr_daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_elec_year <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(elec_yr_pre_data$favorite_count), mean(elec_yr_pre_data$repost_count), mean(elec_yr_daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(elec_yr_post_data$favorite_count), mean(elec_yr_post_data$repost_count), mean(elec_yr_daily_post$n_posts)), 2),
  p_value = c(
    t.test(elec_yr_pre_data$favorite_count, elec_yr_post_data$favorite_count)$p.value,
    t.test(elec_yr_pre_data$repost_count, elec_yr_post_data$repost_count)$p.value,
    t.test(elec_yr_daily_pre$n_posts, elec_yr_daily_post$n_posts)$p.value
  )
)
summary_elec_year$Significant_0.05 <- ifelse(summary_elec_year$p_value < 0.05, "Yes", "No")
```

### Results

```{r}
summary_elec_month$Window <- "Month"
summary_elec_year$Window  <- "Year"
combined_elec <- rbind(summary_elec_month, summary_elec_year)
combined_elec <- combined_elec[, c("Window", "Metric", "Pre_Mean", "Post_Mean", "p_value", "Significant_0.05")]
combined_elec
```

### Plots — Election Loss (Month Window)

```{r, fig.width=8, fig.height=6}
# Combine pre/post month data with a label
elec_month <- bind_rows(
  elec_pre_data %>% mutate(period = "Pre"),
  elec_post_data %>% mutate(period = "Post")
)
elec_month$period <- factor(elec_month$period, levels = c("Pre", "Post"))

# Daily aggregates for the time-series
elec_daily_all <- elec_month %>%
  group_by(day, period) %>%
  summarise(mean_likes = mean(favorite_count),
            mean_retweets = mean(repost_count),
            n_posts = n(), .groups = "drop")

# --- Time-series: Daily Mean Likes ---
p7 <- ggplot(elec_daily_all, aes(x = day, y = mean_likes)) +
  geom_line(color = "steelblue", size = 0.7) +
  geom_point(color = "steelblue", size = 1.2) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Daily Mean Likes Around 2020 Election Loss (Month)",
       x = "Date", y = "Mean Likes") +
  theme_minimal()

# --- Time-series: Daily Mean Retweets ---
p8 <- ggplot(elec_daily_all, aes(x = day, y = mean_retweets)) +
  geom_line(color = "darkorange", size = 0.7) +
  geom_point(color = "darkorange", size = 1.2) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Daily Mean Retweets Around 2020 Election Loss (Month)",
       x = "Date", y = "Mean Retweets") +
  theme_minimal()

# --- Time-series: Daily Post Count ---
p9 <- ggplot(elec_daily_all, aes(x = day, y = n_posts)) +
  geom_line(color = "forestgreen", size = 0.7) +
  geom_point(color = "forestgreen", size = 1.2) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Daily Post Count Around 2020 Election Loss (Month)",
       x = "Date", y = "Number of Posts") +
  theme_minimal()

print(p7)
print(p8)
print(p9)
```

```{r, fig.width=10, fig.height=4}
# --- Boxplots: Pre vs Post ---
bx5 <- ggplot(elec_month, aes(x = period, y = favorite_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Likes: Pre vs Post Election (Month)", x = "", y = "Likes") +
  theme_minimal() + theme(legend.position = "none")

bx6 <- ggplot(elec_month, aes(x = period, y = repost_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Retweets: Pre vs Post Election (Month)", x = "", y = "Retweets") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(bx5, bx6, ncol = 2)
```

---

## Year-Window Comparison (November 3, 2019–2020 vs. November 3, 2020–2021)

We now widen the comparison to the full **year before** vs. **year after** the election loss.

- **Pre-election window:** November 3, 2019 – November 2, 2020
- **Post-election window:** November 3, 2020 – November 2, 2021


### Plots — Election Loss (Year Window)

```{r, fig.width=8, fig.height=6}
# Combine pre/post year data
elec_year <- bind_rows(
  elec_yr_pre_data %>% mutate(period = "Pre"),
  elec_yr_post_data %>% mutate(period = "Post")
)
elec_year$period <- factor(elec_year$period, levels = c("Pre", "Post"))

# Weekly aggregates for cleaner year-long time-series
elec_year$week <- floor_date(elec_year$day, "week")
elec_weekly <- elec_year %>%
  group_by(week) %>%
  summarise(mean_likes = mean(favorite_count),
            mean_retweets = mean(repost_count),
            n_posts = n(), .groups = "drop")

p10 <- ggplot(elec_weekly, aes(x = week, y = mean_likes)) +
  geom_line(color = "steelblue", size = 0.6) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Weekly Mean Likes Around 2020 Election Loss (Year)",
       x = "Date", y = "Mean Likes") +
  theme_minimal()

p11 <- ggplot(elec_weekly, aes(x = week, y = mean_retweets)) +
  geom_line(color = "darkorange", size = 0.6) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Weekly Mean Retweets Around 2020 Election Loss (Year)",
       x = "Date", y = "Mean Retweets") +
  theme_minimal()

p12 <- ggplot(elec_weekly, aes(x = week, y = n_posts)) +
  geom_line(color = "forestgreen", size = 0.6) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Weekly Post Count Around 2020 Election Loss (Year)",
       x = "Date", y = "Number of Posts") +
  theme_minimal()

print(p10)
print(p11)
print(p12)
```

```{r, fig.width=10, fig.height=4}
bx7 <- ggplot(elec_year, aes(x = period, y = favorite_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Likes: Pre vs Post Election (Year)", x = "", y = "Likes") +
  theme_minimal() + theme(legend.position = "none")

bx8 <- ggplot(elec_year, aes(x = period, y = repost_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Retweets: Pre vs Post Election (Year)", x = "", y = "Retweets") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(bx7, bx8, ncol = 2)
```

## Section 4: Post, Pre, and during presidency

```{r}
# Define Era Boundaries
inauguration_date <- as.Date("2017-01-20")
term_end_date     <- as.Date("2021-01-20")

# Create the period factor
term_data <- data %>%
  mutate(period = case_when(
    day < inauguration_date ~ "Pre–Presidency",
    day >= inauguration_date & day <= term_end_date ~ "Presidency",
    day > term_end_date ~ "Post–Presidency"
  )) %>%
  mutate(period = factor(period, levels = c("Post–Presidency", "Pre–Presidency", "Presidency")))

# Generate Summary Table 1
era_summary <- term_data %>%
  group_by(period) %>%
  summarise(
    total_posts = n(),
    mean_likes = mean(favorite_count, na.rm = TRUE),
    median_likes = median(favorite_count, na.rm = TRUE),
    mean_retweets = mean(repost_count, na.rm = TRUE),
    .groups = "drop"
  )

era_summary
```
```{r}
# Filter out the NA row and prepare data for plotting
era_plot_data <- term_data %>%
  filter(!is.na(period)) %>%
  select(period, favorite_count, repost_count) %>%
  tidyr::pivot_longer(cols = c(favorite_count, repost_count), 
                      names_to = "Metric", 
                      values_to = "Count") %>%
  mutate(Metric = ifelse(Metric == "favorite_count", "Likes", "Retweets"))

ggplot(era_plot_data, aes(x = period, y = Count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.1, lwd = 0.5) +
  # Log scale with readable numbers
  scale_y_log10(breaks = c(1, 10, 100, 1000, 10000, 100000, 1000000), 
                labels = scales::label_comma()) +
  facet_wrap(~Metric, scales = "free_y") +
  scale_fill_manual(values = c("Post–Presidency" = "#AEC6CF", 
                               "Pre–Presidency" = "#FFB347", 
                               "Presidency" = "#FF6961")) +
  labs(title = "Engagement Distribution by Political Era",
       subtitle = "Note: Log scale used to compare different orders of magnitude",
       x = "Era", y = "Count (Log Scale)") +
  theme_minimal() +
  theme(legend.position = "none", strip.text = element_text(face = "bold"))
```

## Section 5 - Post-January 6th Analysis 

```{r}
# Define Jan 6th
jan6_date <- as.Date("2021-01-06")

# Create the logical flag
jan_data <- data %>%
  mutate(is_post_jan6 = day >= jan6_date)

# Generate Summary Table 2
jan6_summary <- jan_data %>%
  group_by(is_post_jan6) %>%
  summarise(
    total_posts = n(),
    mean_likes = mean(favorite_count, na.rm = TRUE),
    mean_retweets = mean(repost_count, na.rm = TRUE),
    .groups = "drop"
  )

jan6_summary

```

```{r}
# Prepare Jan 6th data for plotting
jan6_plot_data <- jan_data %>%
  filter(!is.na(is_post_jan6)) %>%
  select(is_post_jan6, favorite_count, repost_count) %>%
  tidyr::pivot_longer(cols = c(favorite_count, repost_count), 
                      names_to = "Metric", 
                      values_to = "Count") %>%
  mutate(Metric = ifelse(Metric == "favorite_count", "Likes", "Retweets"),
         is_post_jan6 = ifelse(is_post_jan6, "Post-Jan 6th", "Pre-Jan 6th"))

ggplot(jan6_plot_data, aes(x = is_post_jan6, y = Count, fill = is_post_jan6)) +
  geom_boxplot(outlier.alpha = 0.1) +
  scale_y_log10(breaks = c(1, 10, 100, 1000, 10000, 100000, 1000000), 
                labels = scales::label_comma()) +
  facet_wrap(~Metric) +
  scale_fill_manual(values = c("Pre-Jan 6th" = "#4E79A7", "Post-Jan 6th" = "#E15759")) +
  labs(title = "Engagement Before vs. After January 6th",
       subtitle = "Comparing the spread of likes and retweets (Log Scale)",
       x = "", y = "Count") +
  theme_minimal() +
  theme(legend.position = "none", strip.text = element_text(face = "bold"))
```


```{r, fig.width=10, fig.height=8, warning=FALSE}
## Section 3 - January 6th Insurrection & Platform Transition

jan6_date       <- as.Date("2021-01-06")
suspension_date <- as.Date("2021-01-08")
ts_return_date  <- as.Date("2022-02-14")

daily_trends <- data %>%
  filter(day >= as.Date("2020-10-01") & day <= as.Date("2022-06-01")) %>%
  group_by(day) %>%
  summarise(
    post_count = n(),
    avg_likes = mean(favorite_count, na.rm = TRUE),
    avg_reposts = mean(repost_count, na.rm = TRUE),
    .groups = "drop"
  )

p_freq <- ggplot(daily_trends, aes(x = day, y = post_count)) +
  annotate("rect", xmin = suspension_date, xmax = ts_return_date, 
           ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  geom_line(color = "#2c3e50", size = 1) +
  # Removed as.numeric() to fix the Date scale warnings
  geom_vline(xintercept = jan6_date, linetype = "dashed", color = "#c0392b", size = 1) +
  scale_y_continuous(limits = c(0, 60), breaks = seq(0, 60, by = 10)) + 
  annotate("label", x = jan6_date + 120, y = 50, label = "Twitter Suspension Period", 
           color = "#c0392b", fontface = "bold", size = 4, fill = "white") +
  labs(title = "Post Frequency & Engagement Surrounding January 6th",
       subtitle = "Activity surge leading to Jan 6th followed by platform hiatus",
       x = NULL, y = "Daily Post Count") +
  theme_minimal(base_size = 14) +
  theme(
    plot.title = element_text(face = "bold", size = 18),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 10, 10, 10)
  )

p_engage <- ggplot(daily_trends, aes(x = day)) +
  annotate("rect", xmin = suspension_date, xmax = ts_return_date, 
           ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "red") +
  geom_line(aes(y = avg_likes, color = "Avg Likes"), size = 1) +
  geom_line(aes(y = avg_reposts, color = "Avg Retweets"), size = 1) +
  # Removed as.numeric() here as well
  geom_vline(xintercept = jan6_date, linetype = "dashed", color = "#c0392b", size = 1) +
  scale_y_log10(
    limits = c(100, 1000000),
    breaks = c(100, 1000, 10000, 50000, 100000, 250000, 500000, 1000000),
    labels = scales::comma_format()
  ) +
  scale_color_manual(values = c("Avg Likes" = "#1a5276", "Avg Retweets" = "#d35400")) +
  labs(x = "Timeline", y = "Engagement Count", color = "Metric") +
  theme_minimal(base_size = 14) +
  theme(
    legend.position = "bottom",
    legend.title = element_blank(),
    panel.grid.minor = element_blank(),
    plot.margin = margin(10, 10, 10, 10)
  )

grid.arrange(p_freq, p_engage, ncol = 1, heights = c(1, 1))
```
## Section 6: Overall trends of posts, retweets, and likes over time


```{r}
start_date <- min(monthly_counts$month_yr)
end_date   <- max(monthly_counts$month_yr)

ggplot(monthly_counts, aes(x = month_yr, y = n_posts)) +
  annotate("rect", xmin = start_date, xmax = inauguration_date, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#FFB347") + # Pre
  annotate("rect", xmin = inauguration_date, xmax = term_end_date, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#FF6961") + # During
  annotate("rect", xmin = term_end_date, xmax = end_date, ymin = -Inf, ymax = Inf, alpha = 0.1, fill = "#AEC6CF") + # Post
  geom_line(color = "#4E79A7", size = 0.7) +
  geom_vline(xintercept = as.numeric(c(inauguration_date, term_end_date)), 
             linetype = "dashed", color = "black", alpha = 0.4) +
  labs(title = "Monthly Post Frequency Over Time",
       subtitle = "Colors indicate: Pre-Presidency (Yellow), Presidency (Red), Post-Presidency (Blue)",
       x = "Year", y = "Number of Posts") +
  theme_minimal()

ggplot(engagement_trends, aes(x = as.Date(month_yr), y = Count)) + 
  annotate("rect", xmin = start_date, xmax = inauguration_date, 
           ymin = 1, ymax = Inf, alpha = 0.15, fill = "#FFB347") +
  annotate("rect", xmin = inauguration_date, xmax = term_end_date, 
           ymin = 1, ymax = Inf, alpha = 0.15, fill = "#FF6961") +
  annotate("rect", xmin = term_end_date, xmax = end_date, 
           ymin = 1, ymax = Inf, alpha = 0.15, fill = "#AEC6CF") +
  
  geom_line(aes(color = Metric), size = 0.8, alpha = 0.8) +
  
  scale_y_log10(breaks = c(10, 100, 1000, 10000, 100000), 
                labels = scales::label_comma(),
                expand = c(0, 0)) + # Removes the gap at the bottom/top
  scale_color_manual(values = c("Likes" = "#E15759", "Retweets" = "#76B7B2")) +
  labs(title = "Engagement Trends (Log Scale)",
       subtitle = "Yellow: Pre-Presidency | Red: Presidency | Blue: Post-Presidency",
       x = "Year", y = "Average Engagement Count") +
  theme_minimal()
```
