---
title: "trump_DataOverTime_EG"
author: "Evan Garcia"
date: "2026-03-01"
output: pdf_document
---

```{r}
library(dplyr)
library(lubridate)
data <- read.csv("filtered_trump_dat.csv")
```


## Section 1 - Candidacy Announcement Analysis (June 16, 2015)

We compare average likes (favorite_count), retweets (repost_count), and daily post counts in the month **before** vs. the month **after** Trump announced his candidacy.

- **Pre-announcement window:** May 16 – June 15, 2015
- **Post-announcement window:** June 16 – July 16, 2015

```{r}
# Parse dates
data$datetime <- ymd_hms(data$date)
data$day <- as.Date(data$datetime)

# Define windows
announce_date <- as.Date("2015-06-16")
pre_start  <- announce_date - 31   # May 16, 2015
pre_end    <- announce_date - 1    # June 15, 2015
post_start <- announce_date        # June 16, 2015
post_end   <- announce_date + 30   # July 16, 2015

# Split data
pre_data  <- data %>% filter(day >= pre_start & day <= pre_end)
post_data <- data %>% filter(day >= post_start & day <= post_end)
```

### Processing 

```{r}
# Compute daily post counts for each window
daily_pre <- pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

daily_post <- post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
all_pre_days  <- data.frame(day = seq(pre_start, pre_end, by = "day"))
all_post_days <- data.frame(day = seq(post_start, post_end, by = "day"))

daily_pre  <- left_join(all_pre_days, daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
daily_post <- left_join(all_post_days, daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_month <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(pre_data$favorite_count), mean(pre_data$repost_count), mean(daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(post_data$favorite_count), mean(post_data$repost_count), mean(daily_post$n_posts)), 2),
  p_value = c(
    t.test(pre_data$favorite_count, post_data$favorite_count)$p.value,
    t.test(pre_data$repost_count, post_data$repost_count)$p.value,
    t.test(daily_pre$n_posts, daily_post$n_posts)$p.value
  )
)
summary_month$Significant_0.05 <- ifelse(summary_month$p_value < 0.05, "Yes", "No")

# compute for year 
yr_pre_start  <- as.Date("2014-06-16")
yr_pre_end    <- as.Date("2015-06-15")
yr_post_start <- as.Date("2015-06-16")
yr_post_end   <- as.Date("2016-06-15")

yr_pre_data  <- data %>% filter(day >= yr_pre_start & day <= yr_pre_end)
yr_post_data <- data %>% filter(day >= yr_post_start & day <= yr_post_end)

yr_daily_pre <- yr_pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

yr_daily_post <- yr_post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
yr_all_pre_days  <- data.frame(day = seq(yr_pre_start, yr_pre_end, by = "day"))
yr_all_post_days <- data.frame(day = seq(yr_post_start, yr_post_end, by = "day"))

yr_daily_pre  <- left_join(yr_all_pre_days, yr_daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
yr_daily_post <- left_join(yr_all_post_days, yr_daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_year <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(yr_pre_data$favorite_count), mean(yr_pre_data$repost_count), mean(yr_daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(yr_post_data$favorite_count), mean(yr_post_data$repost_count), mean(yr_daily_post$n_posts)), 2),
  p_value = c(
    t.test(yr_pre_data$favorite_count, yr_post_data$favorite_count)$p.value,
    t.test(yr_pre_data$repost_count, yr_post_data$repost_count)$p.value,
    t.test(yr_daily_pre$n_posts, yr_daily_post$n_posts)$p.value
  )
)
summary_year$Significant_0.05 <- ifelse(summary_year$p_value < 0.05, "Yes", "No")
```

### Results 

```{r}
summary_month$Window <- "Month"
summary_year$Window  <- "Year"
combined <- rbind(summary_month, summary_year)
combined <- combined[, c("Window", "Metric", "Pre_Mean", "Post_Mean", "p_value", "Significant_0.05")]
combined
```

```{r, fig.width=10, fig.height=4}
# --- Boxplots: Pre vs Post ---
bx1 <- ggplot(cand_month, aes(x = period, y = favorite_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Likes: Pre vs Post Announcement (Month)", x = "", y = "Likes") +
  theme_minimal() + theme(legend.position = "none")

bx2 <- ggplot(cand_month, aes(x = period, y = repost_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Retweets: Pre vs Post Announcement (Month)", x = "", y = "Retweets") +
  theme_minimal() + theme(legend.position = "none")

library(gridExtra)
grid.arrange(bx1, bx2, ncol = 2)
```

---

## Year-Window Comparison (June 16, 2014–2015 vs. June 16, 2015–2016)

We now widen the comparison to the full **year before** vs. **year after** the announcement.

- **Pre-announcement window:** June 16, 2014 – June 15, 2015
- **Post-announcement window:** June 16, 2015 – June 15, 2016


### Plots — Candidacy Announcement (Year Window)

```{r, fig.width=10, fig.height=4}
bx3 <- ggplot(cand_year, aes(x = period, y = favorite_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Likes: Pre vs Post Announcement (Year)", x = "", y = "Likes") +
  theme_minimal() + theme(legend.position = "none")

bx4 <- ggplot(cand_year, aes(x = period, y = repost_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Retweets: Pre vs Post Announcement (Year)", x = "", y = "Retweets") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(bx3, bx4, ncol = 2)
```

---

## Section 2 - Election Loss Analysis (November 3, 2020)

We compare average likes (favorite_count), retweets (repost_count), and daily post counts in the month **before** vs. the month **after** Trump lost the 2020 election.

- **Pre-election window:** October 3 – November 2, 2020
- **Post-election window:** November 3 – December 3, 2020

```{r}
# Define election-loss windows
elec_date <- as.Date("2020-11-03")
elec_pre_start  <- elec_date - 31   # Oct 3, 2020
elec_pre_end    <- elec_date - 1     # Nov 2, 2020
elec_post_start <- elec_date         # Nov 3, 2020
elec_post_end   <- elec_date + 30    # Dec 3, 2020

# Split data
elec_pre_data  <- data %>% filter(day >= elec_pre_start & day <= elec_pre_end)
elec_post_data <- data %>% filter(day >= elec_post_start & day <= elec_post_end)
```

### Processing

```{r}
# Compute daily post counts for each window
elec_daily_pre <- elec_pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

elec_daily_post <- elec_post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
elec_all_pre_days  <- data.frame(day = seq(elec_pre_start, elec_pre_end, by = "day"))
elec_all_post_days <- data.frame(day = seq(elec_post_start, elec_post_end, by = "day"))

elec_daily_pre  <- left_join(elec_all_pre_days, elec_daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
elec_daily_post <- left_join(elec_all_post_days, elec_daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_elec_month <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(elec_pre_data$favorite_count), mean(elec_pre_data$repost_count), mean(elec_daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(elec_post_data$favorite_count), mean(elec_post_data$repost_count), mean(elec_daily_post$n_posts)), 2),
  p_value = c(
    t.test(elec_pre_data$favorite_count, elec_post_data$favorite_count)$p.value,
    t.test(elec_pre_data$repost_count, elec_post_data$repost_count)$p.value,
    t.test(elec_daily_pre$n_posts, elec_daily_post$n_posts)$p.value
  )
)
summary_elec_month$Significant_0.05 <- ifelse(summary_elec_month$p_value < 0.05, "Yes", "No")

# compute for year
elec_yr_pre_start  <- as.Date("2019-11-03")
elec_yr_pre_end    <- as.Date("2020-11-02")
elec_yr_post_start <- as.Date("2020-11-03")
elec_yr_post_end   <- as.Date("2021-11-02")

elec_yr_pre_data  <- data %>% filter(day >= elec_yr_pre_start & day <= elec_yr_pre_end)
elec_yr_post_data <- data %>% filter(day >= elec_yr_post_start & day <= elec_yr_post_end)

elec_yr_daily_pre <- elec_yr_pre_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

elec_yr_daily_post <- elec_yr_post_data %>%
  group_by(day) %>%
  summarise(n_posts = n(), .groups = "drop")

# Fill in zero-post days
elec_yr_all_pre_days  <- data.frame(day = seq(elec_yr_pre_start, elec_yr_pre_end, by = "day"))
elec_yr_all_post_days <- data.frame(day = seq(elec_yr_post_start, elec_yr_post_end, by = "day"))

elec_yr_daily_pre  <- left_join(elec_yr_all_pre_days, elec_yr_daily_pre, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))
elec_yr_daily_post <- left_join(elec_yr_all_post_days, elec_yr_daily_post, by = "day") %>% mutate(n_posts = ifelse(is.na(n_posts), 0, n_posts))


summary_elec_year <- data.frame(
  Metric = c("Likes (favorite_count)", "Retweets (repost_count)", "Daily Post Count"),
  Pre_Mean = round(c(mean(elec_yr_pre_data$favorite_count), mean(elec_yr_pre_data$repost_count), mean(elec_yr_daily_pre$n_posts)), 2),
  Post_Mean = round(c(mean(elec_yr_post_data$favorite_count), mean(elec_yr_post_data$repost_count), mean(elec_yr_daily_post$n_posts)), 2),
  p_value = c(
    t.test(elec_yr_pre_data$favorite_count, elec_yr_post_data$favorite_count)$p.value,
    t.test(elec_yr_pre_data$repost_count, elec_yr_post_data$repost_count)$p.value,
    t.test(elec_yr_daily_pre$n_posts, elec_yr_daily_post$n_posts)$p.value
  )
)
summary_elec_year$Significant_0.05 <- ifelse(summary_elec_year$p_value < 0.05, "Yes", "No")
```

### Results

```{r}
summary_elec_month$Window <- "Month"
summary_elec_year$Window  <- "Year"
combined_elec <- rbind(summary_elec_month, summary_elec_year)
combined_elec <- combined_elec[, c("Window", "Metric", "Pre_Mean", "Post_Mean", "p_value", "Significant_0.05")]
combined_elec
```

### Plots — Election Loss (Month Window)

```{r, fig.width=8, fig.height=6}
# Combine pre/post month data with a label
elec_month <- bind_rows(
  elec_pre_data %>% mutate(period = "Pre"),
  elec_post_data %>% mutate(period = "Post")
)
elec_month$period <- factor(elec_month$period, levels = c("Pre", "Post"))

# Daily aggregates for the time-series
elec_daily_all <- elec_month %>%
  group_by(day, period) %>%
  summarise(mean_likes = mean(favorite_count),
            mean_retweets = mean(repost_count),
            n_posts = n(), .groups = "drop")

# --- Time-series: Daily Mean Likes ---
p7 <- ggplot(elec_daily_all, aes(x = day, y = mean_likes)) +
  geom_line(color = "steelblue", size = 0.7) +
  geom_point(color = "steelblue", size = 1.2) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Daily Mean Likes Around 2020 Election Loss (Month)",
       x = "Date", y = "Mean Likes") +
  theme_minimal()

# --- Time-series: Daily Mean Retweets ---
p8 <- ggplot(elec_daily_all, aes(x = day, y = mean_retweets)) +
  geom_line(color = "darkorange", size = 0.7) +
  geom_point(color = "darkorange", size = 1.2) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Daily Mean Retweets Around 2020 Election Loss (Month)",
       x = "Date", y = "Mean Retweets") +
  theme_minimal()

# --- Time-series: Daily Post Count ---
p9 <- ggplot(elec_daily_all, aes(x = day, y = n_posts)) +
  geom_line(color = "forestgreen", size = 0.7) +
  geom_point(color = "forestgreen", size = 1.2) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Daily Post Count Around 2020 Election Loss (Month)",
       x = "Date", y = "Number of Posts") +
  theme_minimal()

print(p7)
print(p8)
print(p9)
```

```{r, fig.width=10, fig.height=4}
# --- Boxplots: Pre vs Post ---
bx5 <- ggplot(elec_month, aes(x = period, y = favorite_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Likes: Pre vs Post Election (Month)", x = "", y = "Likes") +
  theme_minimal() + theme(legend.position = "none")

bx6 <- ggplot(elec_month, aes(x = period, y = repost_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Retweets: Pre vs Post Election (Month)", x = "", y = "Retweets") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(bx5, bx6, ncol = 2)
```

---

## Year-Window Comparison (November 3, 2019–2020 vs. November 3, 2020–2021)

We now widen the comparison to the full **year before** vs. **year after** the election loss.

- **Pre-election window:** November 3, 2019 – November 2, 2020
- **Post-election window:** November 3, 2020 – November 2, 2021


### Plots — Election Loss (Year Window)

```{r, fig.width=8, fig.height=6}
# Combine pre/post year data
elec_year <- bind_rows(
  elec_yr_pre_data %>% mutate(period = "Pre"),
  elec_yr_post_data %>% mutate(period = "Post")
)
elec_year$period <- factor(elec_year$period, levels = c("Pre", "Post"))

# Weekly aggregates for cleaner year-long time-series
elec_year$week <- floor_date(elec_year$day, "week")
elec_weekly <- elec_year %>%
  group_by(week) %>%
  summarise(mean_likes = mean(favorite_count),
            mean_retweets = mean(repost_count),
            n_posts = n(), .groups = "drop")

p10 <- ggplot(elec_weekly, aes(x = week, y = mean_likes)) +
  geom_line(color = "steelblue", size = 0.6) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Weekly Mean Likes Around 2020 Election Loss (Year)",
       x = "Date", y = "Mean Likes") +
  theme_minimal()

p11 <- ggplot(elec_weekly, aes(x = week, y = mean_retweets)) +
  geom_line(color = "darkorange", size = 0.6) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Weekly Mean Retweets Around 2020 Election Loss (Year)",
       x = "Date", y = "Mean Retweets") +
  theme_minimal()

p12 <- ggplot(elec_weekly, aes(x = week, y = n_posts)) +
  geom_line(color = "forestgreen", size = 0.6) +
  geom_vline(xintercept = as.numeric(elec_date), linetype = "dashed", color = "red", size = 0.8) +
  annotate("text", x = elec_date, y = Inf, label = "Election Day", vjust = 2, hjust = -0.1, color = "red", size = 3.5) +
  labs(title = "Weekly Post Count Around 2020 Election Loss (Year)",
       x = "Date", y = "Number of Posts") +
  theme_minimal()

print(p10)
print(p11)
print(p12)
```

```{r, fig.width=10, fig.height=4}
bx7 <- ggplot(elec_year, aes(x = period, y = favorite_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Likes: Pre vs Post Election (Year)", x = "", y = "Likes") +
  theme_minimal() + theme(legend.position = "none")

bx8 <- ggplot(elec_year, aes(x = period, y = repost_count, fill = period)) +
  geom_boxplot(outlier.alpha = 0.3) +
  scale_fill_manual(values = c("Pre" = "#AEC6CF", "Post" = "#FFB347")) +
  labs(title = "Retweets: Pre vs Post Election (Year)", x = "", y = "Retweets") +
  theme_minimal() + theme(legend.position = "none")

grid.arrange(bx7, bx8, ncol = 2)
```

