---
title: "Untitled"
author: "Prem Kollu"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(stringr)
trump_dat <- read_csv("filtered_trump_dat.csv")
```
```{r}
glimpse(trump_dat)
head(trump_dat$text, 3)
```

```{r}
# ============================================================
# Feature Engineering: Exclamation Points, Capital Letters,
# Hashtags, Total Letters
# ============================================================

trump_features <- trump_dat %>%
  mutate(
    exclamation_count = str_count(text, fixed("!")),
    capital_count     = str_count(text, "[A-Z]"),
    hashtag_count     = str_count(text, "#\\S+"),
    total_letters     = str_count(text, "[A-Za-z]")
  )

# Quick check
trump_features %>%
  select(text, exclamation_count, capital_count, hashtag_count, total_letters) %>%
  head(10)
```

```{r}
# Summary stats across all 58,020 tweets
trump_features %>%
  select(exclamation_count, capital_count, hashtag_count, total_letters) %>%
  summary()
```
```{r}
# A visualization
library(ggplot2)
library(gridExtra)

p1 <- ggplot(trump_features, aes(x = exclamation_count)) +
  geom_histogram(binwidth = 1, fill = "steelblue", color = "white") +
  coord_cartesian(xlim = c(0, 10)) +
  labs(title = "Exclamation Points per Tweet", x = "Count", y = "# of Tweets") +
  theme_minimal()

p2 <- ggplot(trump_features, aes(x = capital_count)) +
  geom_histogram(binwidth = 2, fill = "tomato", color = "white") +
  coord_cartesian(xlim = c(0, 100)) +
  labs(title = "Capital Letters per Tweet", x = "Count", y = "# of Tweets") +
  theme_minimal()

p3 <- ggplot(trump_features, aes(x = hashtag_count)) +
  geom_histogram(binwidth = 1, fill = "seagreen", color = "white") +
  coord_cartesian(xlim = c(0, 8)) +
  labs(title = "Hashtags per Tweet", x = "Count", y = "# of Tweets") +
  theme_minimal()

p4 <- ggplot(trump_features, aes(x = total_letters)) +
  geom_histogram(binwidth = 10, fill = "purple", color = "white") +
  coord_cartesian(xlim = c(0, 500)) +
  labs(title = "Total Letters per Tweet", x = "Count", y = "# of Tweets") +
  theme_minimal()

grid.arrange(p1, p2, p3, p4, ncol = 2)
```
Exclamation Points — most tweets have 0-1, with a sharp drop after that. He uses them, but not excessively on most tweets.

Capital Letters — nice bell-ish shape peaking around 10-15 capitals per tweet. This is pretty normal for any English text.

Hashtags — the vast majority of tweets have 0 hashtags. Trump barely uses them, which is interesting given how hashtag-heavy political Twitter typically is.

Total Letters — interesting bimodal shape — there's a peak around 100 letters and another around 280. This likely reflects the old 140-character limit vs the new 280-character limit Twitter introduced in 2017. That's a really cool finding worth mentioning in your report.

